visitation_bias

### 2021-01-22 code/visitation_bias.Rmd
backup

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(mgcv)
library(mgcViz)
library(brms)
library(lme4)
library(bayesplot)
library(vegan)
library(tidyverse)
library(lubridate)
library(tidymodels)
```

@@

@@

@@

# 1. Quantifying bias
How does overall bias (Blüthgen's d') vary across species and through elev*time? For now, we will focus on bias at the level of plant species. We'll start considering traits later on when our goal is to explain patterns of bias.
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.sp", n())))

# Grouped by k.type.ss
bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("ktype", n())))

# Grouped by plant.family
bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
  left_join(site_data) %>%
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  mutate(criterion = factor(rep("plant.family", n())))

# Collated
bias <- bind_rows(bias_plantsp, 
                  bias_ktype, 
                  bias_family) %>%
  mutate(bb.sp = factor(bb.sp),
         yday = yday(date),
         elev = elev.mean/10) 
```

## 1.2 Visualize
```{r}
# Compare mean bias by criterion
ggplot(bias, aes(criterion, dprime)) +
  geom_boxplot()

# Bias ~ elevation * criterion
ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = TRUE)

ggplot(bias, aes(elev.mean, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime, color = criterion)) +
  geom_point(alpha = 0.25) +
  geom_smooth(se = FALSE, method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# Bias ~ year; Blüthgen's d' seems very robust to sampling effects, since it does not vary across years
ggplot(bias, aes(year, dprime, color = criterion)) +
  geom_boxplot()

ggplot(bias, aes(year, dprime, color = criterion)) +
  geom_boxplot() +
  facet_wrap(~bb.sp)

# Does bias vary with abundance?
ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(bb.sp.abund, dprime, color = criterion)) +
  #geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp, )

ggplot(bias, aes(log(bb.sp.abund), dprime, color = criterion)) +
  geom_point(alpha = 0.1) +
  geom_smooth()

ggplot(bias, aes(log(bb.sp.prop.abund), dprime)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~bb.sp)
```

## 1.3 Multi-level model: bias ~ yday * elevation + (1 + elevation | bb.sp)
```{r}
# I am interested both in the variation of bias through elevation and time and in the intercept effect of each bumble bee species

# First, let's look at the distribution of bias per BB sp
ggplot(filter(bias, criterion == "plant.sp"), aes(dprime)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Now, let's try to get a sense for the shape of the response to elevation and time; we might expect to find nonlinearities here
ggplot(filter(bias, criterion == "plant.sp"), aes(elev, dprime)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(filter(bias, criterion == "plant.sp"), aes(yday, dprime)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# After inspecting these plots, I see no reason to indulge in wiggly models.

# This model includes a global interaction term for yday and elevation, with a varying intercept for bb.sp and a varying slow for elev | bb.sp. I think this should give me all the estimation I'm interested in. It's tricky to make it converge, though.
brm.plantsp11 <- brm(dprime ~ yday * elev + (1 + elev | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 4000,
                  #family = "skew_normal",
                  file = "../output/brm_plantsp11",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp11)
prior_summary(brm.plantsp11)
plot(brm.plantsp11)
parnames(brm.plantsp11)
mcmc_plot(brm.plantsp11, pars = c("^r_"))
mcmc_plot(brm.plantsp11, pars = c("elev"))
mcmc_plot(brm.plantsp11, pars = c("^b_"))
mcmc_plot(brm.plantsp11)
conditional_effects(brm.plantsp11)
pp_check(brm.plantsp11)
bayes_R2(brm.plantsp11)
hypothesis(brm.plantsp11, "b_elev = 0", class = NULL)

# Alternatively, we can drop the varying slope term to simplify the model a bit. Still doesn't quite converge.
brm.plantsp1 <- brm(dprime ~ yday * elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp1",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp1)
prior_summary(brm.plantsp1)
plot(brm.plantsp1)
parnames(brm.plantsp1)
mcmc_plot(brm.plantsp1, pars = c("^r_"))
mcmc_plot(brm.plantsp1, pars = c("elev"))
mcmc_plot(brm.plantsp1, pars = c("^b_"))
mcmc_plot(brm.plantsp1)
conditional_effects(brm.plantsp1)
pp_check(brm.plantsp1)
bayes_R2(brm.plantsp1)
hypothesis(brm.plantsp1, "b_elev = 0", class = NULL)

# We can simplify the model still more by dropping the yday*elev interaction term. Now it converges. But I think the takeaway is that neither elevation noy yday give us a clear signal with respect  to dprime.
brm.plantsp2 <- brm(dprime ~ yday + elev + (1 | bb.sp), 
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp2",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp2)
prior_summary(brm.plantsp2)
plot(brm.plantsp2)
parnames(brm.plantsp2)
mcmc_plot(brm.plantsp2, pars = c("^r_"))
mcmc_plot(brm.plantsp2, pars = c("elev"))
mcmc_plot(brm.plantsp2, pars = c("^b_"))
mcmc_plot(brm.plantsp2)
conditional_effects(brm.plantsp2)
pp_check(brm.plantsp2)
bayes_R2(brm.plantsp2)
hypothesis(brm.plantsp2, "b_elev = 0", class = NULL)

# This, I think, would also be a legitimate way to specify the model, but it would not tell us anything about elevation or time. It's the simplest, and that's nice. Well-behaved. Captures about twice as much variance as the previous model, and it wins the WAIC context easily.
brm.plantsp0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp0",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp0)
plot(brm.plantsp0)
parnames(brm.plantsp0)

# I guess the message here is that both bb.sp and site contribute to the variation of dprime, but site more so. Biologically, I think the meaning of prime variation by site is that some sites have greater relative abundance of *attractive* flora than others. A site with loads of something that is never visited and a little bit of clover will yield a high dprime, while a site with a high abundance of attractive floral would yield a low dprime.
mcmc_plot(brm.plantsp0, pars = c("^r_"))
mcmc_plot(brm.plantsp0, pars = c("^r_bb.sp")) 
ggsave("../output/brm.plantsp0_plot1.png")
mcmc_plot(brm.plantsp0, pars = c("^b_"))
pp_check(brm.plantsp0)
bayes_R2(brm.plantsp0)

# Another option would be to drop just elevation from the model. This runs clean. Doesn't explain much more variance than the simpler model, though.
brm.plantsp00 <- brm(dprime ~ yday + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  file = "../output/brm_plantsp00",
                  data = filter(bias, criterion ==  "plant.sp"))

summary(brm.plantsp00)
plot(brm.plantsp00)
parnames(brm.plantsp00)
mcmc_plot(brm.plantsp00, pars = c("^r_"))
mcmc_plot(brm.plantsp00, pars = c("^b_"))
conditional_effects(brm.plantsp00)
pp_check(brm.plantsp00)
bayes_R2(brm.plantsp00)

# Let's do a formal comparison of the last three models
waic(brm.plantsp2, brm.plantsp0, brm.plantsp00)

# The one including yday is a *slight* winner, but the margin is so small as to be meaningless. I think the simplest model is the one we want (brm.plantsp0).

# Let's try the same model now, but for ktype.
brm.ktype0 <- brm(dprime ~ 1 + (1 | bb.sp) + (1 | site),
                  control = list(max_treedepth = 10),
                  cores = 4,
                  iter = 2000,
                  family = "skew_normal",
                  file = "../output/brm_ktype0",
                  data = filter(bias, criterion ==  "ktype"))

summary(brm.ktype0)
plot(brm.ktype0)
parnames(brm.ktype0)
mcmc_plot(brm.ktype0, pars = c("^r_"))
mcmc_plot(brm.ktype0, pars = c("^b_"))
conditional_effects(brm.ktype0)
pp_check(brm.ktype0)
bayes_R2(brm.ktype0)
```

## 1.4 Define hypotheses
```{r}
# Is d' different from 0?
H1 <- paste("Intercept = 0")

# Are some species more biased in their foraging than others? 
H.bss_gers <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[gers,Intercept]")
H.bss_hort <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hort,Intercept]")
H.bss_humi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[humi,Intercept]")
H.bss_hypn <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[hypn,Intercept]")
H.bss_jone <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[jone,Intercept]")
H.bss_lapi <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[lapi,Intercept]")
H.bss_mend <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mend,Intercept]")
H.bss_mont <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[mont,Intercept]")
H.bss_muci <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[muci,Intercept]")
H.bss_pasc <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pasc,Intercept]")
H.bss_prat <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[prat,Intercept]")
H.bss_psit <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[psit,Intercept]")
H.bss_pyre <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[pyre,Intercept]")
H.bss_soro <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[soro,Intercept]")
H.bss_wurf <- paste("r_bb.sp[bss,Intercept] = r_bb.sp[wurf,Intercept]")

H.gers_hort <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hort,Intercept]")
H.gers_humi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[humi,Intercept]")
H.gers_hypn <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[hypn,Intercept]")
H.gers_jone <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[jone,Intercept]")
H.gers_lapi <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[lapi,Intercept]")
H.gers_mend <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mend,Intercept]")
H.gers_mont <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[mont,Intercept]")
H.gers_muci <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[muci,Intercept]")
H.gers_pasc <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pasc,Intercept]")
H.gers_prat <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[prat,Intercept]")
H.gers_psit <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[psit,Intercept]")
H.gers_pyre <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[pyre,Intercept]")
H.gers_soro <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[soro,Intercept]")
H.gers_wurf <- paste("r_bb.sp[gers,Intercept] = r_bb.sp[wurf,Intercept]")

H.hort_humi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[humi,Intercept]")
H.hort_hypn <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[hypn,Intercept]")
H.hort_jone <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[jone,Intercept]")
H.hort_lapi <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[lapi,Intercept]")
H.hort_mend <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mend,Intercept]")
H.hort_mont <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[mont,Intercept]")
H.hort_muci <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[muci,Intercept]")
H.hort_pasc <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pasc,Intercept]")
H.hort_prat <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[prat,Intercept]")
H.hort_psit <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[psit,Intercept]")
H.hort_pyre <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[pyre,Intercept]")
H.hort_soro <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[soro,Intercept]")
H.hort_wurf <- paste("r_bb.sp[hort,Intercept] = r_bb.sp[wurf,Intercept]")

H.humi_hypn <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[hypn,Intercept]")
H.humi_jone <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[jone,Intercept]")
H.humi_lapi <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[lapi,Intercept]")
H.humi_mend <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mend,Intercept]")
H.humi_mont <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[mont,Intercept]")
H.humi_muci <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[muci,Intercept]")
H.humi_pasc <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pasc,Intercept]")
H.humi_prat <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[prat,Intercept]")
H.humi_psit <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[psit,Intercept]")
H.humi_pyre <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[pyre,Intercept]")
H.humi_soro <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[soro,Intercept]")
H.humi_wurf <- paste("r_bb.sp[humi,Intercept] = r_bb.sp[wurf,Intercept]")

>>>>

---
title: "visitation_bias"
author: "Doug Sponsler"
date: "12/14/2020"
output: html_document
---
# Preparations
## Packages 
```{r, warning=FALSE, include=FALSE, echo = FALSE}
library(bipartite)
library(brms)
library(lme4)
library(sjPlot)
library(DHARMa)
library(bayesplot)
library(tidybayes)
library(modelr)
library(ggdist)
library(vegan)
library(ggvegan)
library(ggrepel)
library(treemapify)
library(patchwork)
library(tidymodels)
library(tidyverse)
library(lubridate)

rstan::rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
```

@@

ggplot(bb_traits, aes(pbl.w, pbl.w.durieaux,
                      color = bb.sp, pch = factor(pbl.w.class2))) +
  geom_point()

@@

ktype_families <- visit_cover %>%
  select(plant.sp, plant.family, k.type.ss) %>%
  distinct()

ggplot(ktype_families, aes(plant.family, fill = k.type.ss)) +
  geom_bar(position = "fill") +
  theme(axis.text.x = element_text(angle = 90))

@@

# 1. Quantifying bias
How does overall bias (Blüthgen's d') vary across species and through elev*time? For now, we will focus on bias at the level of plant species. We'll start considering traits later on when our goal is to explain patterns of bias.
## 1.1 Calculate Blüthgen's d' (i.e. visitation bias)
```{r}
# No grouping -- just plant.sp
bias_plantsp <- biaseval(visit_cover, plant.sp, site, date) %>%
  rename(dprime.sp = dprime) %>%
  select(site, date, bb.sp, dprime.sp)

# Grouped by k.type.ss
bias_ktype <- biaseval(visit_cover, k.type.ss, site, date) %>%
  rename(dprime.ktype = dprime) %>%
  select(site, date, bb.sp, dprime.ktype)

# Grouped by plant.family
bias_family <- biaseval(visit_cover, plant.family, site, date) %>%
  rename(dprime.fam = dprime) %>%
  select(site, date, bb.sp, dprime.fam)

# Collated
bias <- full_join(bias_plantsp, bias_ktype) %>%
  full_join(bias_family) %>%
  left_join(site_data) %>% # add site data
  left_join(bb_abundance) %>% # add BB abundance data
  left_join(fl_abundance) %>% # add fl abundance
  filter(bb.sp.abund >= 3) %>% # dprime tends to become binary when n is very low; this threshold removes more than half the data points, but almost all those were ones and zeros. These ones and zeros probably don't mean much, and they make the models hard to fit.
  mutate(bb.sp = factor(bb.sp), 
         yday = yday(date),
         elev = elev.mean/10) %>% #  scale this down for convenience
  mutate(yday0 = scale(yday),  # normalized versions of variables
         elev0 = scale(elev),
         dprime.sp0 = scale(dprime.sp),
         dprime.ktype0 = scale(dprime.ktype),
         dprime.fam0 = scale(dprime.fam)) %>%
  select(site, date, bb.sp, dprime.sp, dprime.sp0, dprime.ktype, 
         dprime.ktype0, dprime.fam, dprime.fam0, elev, elev0, 
         yday, yday0, bb.sp.abund, bb.sp.prop.abund, total.flower.cover)
```

## 1.3 Multi-level model: bias ~ yday * elevation + (1 + elevation | bb.sp)
### 1.3.1 Preliminary visualization
```{r}
# I am interested both in the variation of bias through elevation and time and in the intercept effect of each bumble bee species

ggplot(bias, aes(dprime.sp)) +
  geom_histogram()

# First, let's look at the distribution of bias per BB sp
ggplot(bias, aes(dprime.sp)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(dprime.ktype)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

ggplot(bias, aes(dprime.fam)) +
  geom_histogram() +
  facet_wrap(~bb.sp)

# Now, let's try to get a sense for the shape of the response to elevation and time; we might expect to find nonlinearities here
ggplot(bias, aes(elev, dprime.sp)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.sp)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(elev, dprime.ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.ktype)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(elev, dprime.fam)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

ggplot(bias, aes(yday, dprime.fam)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_wrap(~bb.sp)

# After inspecting these plots, I see no reason to indulge in wiggly models.
```

### 1.3.3 brms models
There's a lot I don't understand about the standardization of variables in the context of Bayesian MLM, but for now I have opted to scale and center the predictor variables but not the response. My reason for scaling and centering the predictors is that is massively speeds up estimation and enables the interaction model to converge (it would not converge with raw variables). My reason for not scaling the response is that it doesn't change the outcome of the model (I've checked) and I am interested in seeing the global intercept for d' on its native scale so that I can compare it with previously reported values. Right now, I am using all default priors. This means flat priors for global effects and student priors for varying effects (I think). My understanding is that brms internally scales priors to match the scale of the data. If I want to define my own priors, which I know I really should, particularly to replace the flat priors on the global effects, then I need to understand how the scaling of the variables affects the specification of the prior.
```{r}
# elev*yday interaction model
brm_dprime.sp00 <- brm(dprime.sp ~ elev0 * yday0 + (1 | bb.sp),
                       #+ (1 | site),
                       cores = 4,
                       iter = 2000,
                       family = "skew_normal", # this accounts for a slight left-skew in the posterior
                       file = "../output/brm_dprime.sp00",
                       data = bias)

summary(brm_dprime.sp00)
prior_summary(brm_dprime.sp00)
parnames(brm_dprime.sp00)
pp_check(brm_dprime.sp00)
bayes_R2(brm_dprime.sp00)

# Intercept-only model
brm_dprime.sp01 <- brm(dprime.sp ~ 1 + (1 | bb.sp) + (1 | site),
                       cores = 4,
                       iter = 2000,
                       family = "skew_normal",
                       file = "../output/brm_dprime.sp01",
                       data = bias)

summary(brm_dprime.sp01)
prior_summary(brm_dprime.sp01)
parnames(brm_dprime.sp01)
pp_check(brm_dprime.sp01)
bayes_R2(brm_dprime.sp01)

# Compare interaction model and intercept-only model. This favors the intercept-only model.
loo(brm_dprime.sp00, brm_dprime.sp01)
```

#### 1.3.3.1 Plot
```{r}
get_variables(brm_dprime.sp01)

brm_dprime.sp01_plotframe <- brm_dprime.sp01 %>%
  spread_draws(b_Intercept, r_bb.sp[condition, term]) %>% 
  mutate(condition_mean = b_Intercept + r_bb.sp)

brm_dprime.sp01_plotframe2 <- brm_dprime.sp01 %>%
  spread_draws(b_Intercept, r_site[condition, term]) %>% 
  mutate(condition_mean = b_Intercept + r_site)

brm_dprime.sp01_subplot1 <- ggplot(brm_dprime.sp01_plotframe, aes(condition_mean, condition)) +
  stat_halfeye() +
  ylab("Bumble bee species") +
  xlab("Discrimination (d')") +
  theme_light()

brm_dprime.sp01_subplot2 <- ggplot(brm_dprime.sp01_plotframe2, aes(condition_mean, condition)) +
  stat_halfeye() +
  ylab("Site") +
  xlab("Discrimination (d')") +
  theme_light()

brm_dprime.sp01_plot1 <- brm_dprime.sp01_subplot1 | brm_dprime.sp01_subplot2

ggsave(brm_dprime.sp01_plot1, file = "../output/brm_dprime.sp01_plot1.png")
```

# 2. Explaining bias
## 2.1 Calculate visitation probability per plant speces/ktype/family
```{r}
pvis_ktype <- visit_cover %>%
  group_by(site, date, bb.sp, plant.sp) %>%
  summarize(flower.cover = sum(flower.cover), # get raw flower cover and visits by site*date*bb.sp*plant.sp
            visits = sum(visits)) %>%
  group_by(site, date) %>%
  mutate(prop.flower.cover = flower.cover/sum(flower.cover), # proportional flower cover
         prop.visits = visits/sum(visits), # proportional visits
         visits.bin = factor(if_else(visits > 0, 1, 0))) %>% # binary visits
  group_by(site, date, bb.sp) %>%
  mutate(bb.sp.abund = sum(visits)) %>% # abundance of each BB sp. per site-date
  ungroup() %>%
  left_join(fl_traits) %>% # add floral traits
  left_join(bb_traits) %>% # add BB traits
  mutate(k.type.ss = factor(k.type.ss), # recode variables
         pbl.w.class2 = factor(pbl.w.class2),
         visits = as.integer(visits)) %>%
  filter(!k.type.ss %in% c("9", "10")) %>% # Drop ktypes 9 and 10 since both are extremely rare
  left_join(site_data)  %>%
  mutate(prop.flower.cover0 = scale(prop.flower.cover), # normalized predictors
         bb.sp.abund0 = scale(bb.sp.abund),
         elev.mean0 = scale(elev.mean),
         yday = yday(date), # and add yday variable
         yday0 = scale(yday)) %>%
  select(site, elev.mean0, elev.mean, elev.class2, date, yday, yday0, bb.sp, plant.sp, visits, 
         visits.bin, prop.visits, prop.flower.cover, prop.flower.cover0,
         k.type.ss, bb.sp.abund, bb.sp.abund0, pbl.w.class2) 
```

## 2.2 Multi-level models

### 2.2.1 Preliminary visualization
The tons of zeros are due to the lots of plant species that were present at very low abundance and received few or zero visits. This might not be ideal. We could opt for one of our alternative bias measure, like fold change, which all give something like a poisson distribution, but I don't know if this would be an improvement. We could also filter our data to include only plants that were above some threshold of relative abundance, say, 1%.
```{r}
# Visit.prop
ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(prop.visits)) +
  geom_histogram()

# Visits.bin
ggplot(pvis_ktype, aes(k.type.ss, fill = visits.bin)) +
  geom_bar()

ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth() +
  facet_grid(bb.sp~k.type.ss)

# Visits.bin filtered
ggplot(filter(pvis_ktype, prop.flower.cover >= 0.05), aes(k.type.ss, fill = visits.bin)) +
  geom_bar()
```

### 2.2.2 brms models
```{r}
# Poisson models of raw visit counts
brm_pvis00 <- brm(visits ~ prop.flower.cover0 + 
                    bb.sp.abund0 + 
                    (1 | k.type.ss) + 
                    (1 | site) + 
                    (1 | bb.sp),
                  #control = list(adapt_delta = 0.99),
                  family = "poisson",
                  file = "./output/brm_pvis00b",
                  chains = 4,
                  #cores = 4,
                  iter = 2000,
                  data = pvis_ktype)

summary(brm_pvis00)
prior_summary(brm_pvis00)
parnames(brm_pvis00)
mcmc_plot(brm_pvis00, pars = c("^r_k"))
mcmc_plot(brm_pvis00, pars = c("^r_s"))
mcmc_plot(brm_pvis00, pars = c("^b_"))
mcmc_plot(brm_pvis00)
conditional_effects(brm_pvis00)
pp_check(brm_pvis00)
bayes_R2(brm_pvis00)

# Perhaps I should add the bb.sp varying intercept effect; otherwise, I think I could be accused of pseudo replication.
brm_pvis11 <- brm(visits ~ prop.flower.cover0 + 
                    bb.sp.abund0 + 
                    (1 | k.type.ss:pbl.w.class2) + 
                    (1 | site) +
                    (1 | bb.sp),
                  family = "poisson",
                  file = "../output/brm_pvis11",
                  chains = 4,
                  #cores = 4,
                  iter = 2000,
                  data = pvis_ktype)

summary(brm_pvis11)
prior_summary(brm_pvis11)
parnames(brm_pvis11)
mcmc_plot(brm_pvis11, pars = c("^r_k"))
mcmc_plot(brm_pvis11, pars = c("^r_b"))
mcmc_plot(brm_pvis11, pars = c("^b_"))
mcmc_plot(brm_pvis11)
conditional_effects(brm_pvis11)
pp_check(brm_pvis11)
bayes_R2(brm_pvis11)
```

#### 2.2.2.1 Plot
```{r}
# Note that the intercept is on the log scale, so if you want to transform it back to predicted # visits, you have to exponentiate it. I think.

## Interaction model
brm_pvis11_plotframe <- brm_pvis11 %>%
  spread_draws(b_Intercept, `r_k.type.ss:pbl.w.class2`[condition, term]) %>% # the back ticks are necessary to escape the :
  rename(r_ktype.pbl = `r_k.type.ss:pbl.w.class2`) %>%
  mutate(condition_mean = b_Intercept + r_ktype.pbl) %>%
  separate(condition, c("ktype", "pbl.class"), sep = "_", remove = FALSE) %>%
  mutate(pbl.class = case_when(
    pbl.class == "1" ~ "short",
    pbl.class == "2" ~ "med",
    pbl.class == "3" ~ "long")
    ) %>%
  mutate(ktype = case_when(
    ktype == "0" ~ "NA (0)",
    ktype == "1" ~ "Disc (1)",
    ktype == "2" ~ "Funnel (2)",
    ktype == "3" ~ "Bell (3)",
    ktype == "4" ~ "Stalk-disc (4)",
    ktype == "5" ~ "Lip (5)",
    ktype == "6" ~ "Flag (6)",
    ktype == "7" ~ "Head (7)")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc (1)", "Funnel (2)", "Bell (3)", "Stalk-disc (4)",
                                   "Lip (5)", "Flag (6)", "Head (7)", "NA (0)")))

# ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
#   stat_pointinterval() +
#   #stat_halfeye(alpha = 0.5, interval_alpha = 1, point_alpha = 1, size = 1, .width = 0.95) +
#   facet_grid(ktype ~ .) +
#   theme_light() +
#   labs(color = "Tongue length", x = "Visits (log-scale)") +
#   guides(fill = FALSE) +
#   theme(axis.title.y=element_blank(),
#         axis.text.y=element_blank(),
#         axis.ticks.y=element_blank())

ggplot(brm_pvis11_plotframe, aes(condition_mean, pbl.class, color = pbl.class, fill = pbl.class)) +
  #stat_pointinterval() +
  stat_halfeye(alpha = 0.4, interval_alpha = 1, point_alpha = 1) +
  facet_wrap(~ktype, scales = "free") +
  theme_light() +
  labs(color = "Tongue length", x = "Visits (log-scale)") +
  guides(fill = FALSE) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())

ggsave(file = "../output/brm_pvis11_plot1.png", width = 7, height = 5.5)

# Just ktype model
brm_pvis00_plotframe <- brm_pvis00 %>%
  spread_draws(b_Intercept, r_k.type.ss[condition, term]) %>% # the back ticks are necessary to escape the :
  mutate(condition_mean = b_Intercept + r_k.type.ss) %>%
  mutate(ktype = case_when(
    condition == "0" ~ "NA",
    condition == "1" ~ "Disc",
    condition == "2" ~ "Funnel",
    condition == "3" ~ "Bell",
    condition == "4" ~ "Stalk-disc",
    condition == "5" ~ "Lip",
    condition == "6" ~ "Flag",
    condition == "7" ~ "Head")
  ) %>%
  mutate(ktype = factor(ktype,
                        levels = c("Disc", "Funnel", "Bell", "Stalk-disc",
                                   "Lip", "Flag", "Head", "NA")))

ggplot(brm_pvis00_plotframe, aes(condition_mean, ktype)) +
  stat_halfeye() +
  theme_light() +
  guides(fill = FALSE) +
  xlab("Visits (log-scale)") +
  ylab("Morphotype")

ggsave(file = "../output/brm_pvis00_plot1.png", width = 7, height = 5.5)
```

### 2.2.3 The lability of bias: How does the foraging bias of each species with respect to each morphotype vary through elevation and time?
```{r}
ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

ggplot(pvis_ktype, aes(elev.mean, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

ggplot(pvis_ktype, aes(yday0, prop.visits)) +
  geom_point() +
  geom_smooth(method = "gam", formula = y ~ s(x, k = 5)) +
  facet_grid(bb.sp ~ k.type.ss)

brm_pviselev00 <- brm(visits ~ prop.flower.cover0 + 
                        bb.sp.abund0 + 
                        (1 + elev.mean0:yday0 | k.type.ss:bb.sp),
                      family = "poisson",
                      file = "../output/brm_pviselev00",
                      chains = 4,
                      cores = 4,
                      iter = 2000,
                      data = pvis_ktype)

summary(brm_pviselev00)
prior_summary(brm_pviselev00)
parnames(brm_pviselev00)
mcmc_plot(brm_pviselev00, pars = c("^r_k"))
mcmc_plot(brm_pviselev00, pars = c("^r_s"))
mcmc_plot(brm_pviselev00, pars = c("^b_"))
mcmc_plot(brm_pviselev00)
conditional_effects(brm_pviselev00)
pp_check(brm_pviselev00)
bayes_R2(brm_pviselev00)
```
